How cryptic is cryptic diversity? Machine learning approaches to plastral variation in \textit{Emys marmorata}.

### 2013-07-01 documents/turtle_paper.Rnw
methods. pam.

\subsection{Machine learning analyses}
\subsubsection{Unsupervised learning}
% PAM
%   Riemannian shape distance
%   gap statistic
PAM was conducted using the \texttt{cluster} package for R \citep{Maechler2013}.

@@

% model training
%   tuning parameters
%     grid search
%       what are the tuning parameters for each model?
%     best AUC ROC
In order to prevent over fitting each machine learning model, tuning parameters were estimated using 10-fold cross-validation (CV) across a grid search of all tuning parameter combinations. Optimal tuning parameter values were selected based on area under the receiver operating characteristic curve (AUC ROC). Multiclass AUC ROC was estimated using the all-against one strategy described in \citet{Hand2001}.

%   model selection
%     AICc
For the multinomial logistic regression models, PCs were added sequentially in order to increase the overall amount of variation in shape included in each model and the final model was that with the lowest AICc \cite{Burnham2002a}. This procedure was used because the optimal number of PCs to include is unknown, and while including all of the PCs of shape would mean that all of the variability in plastron shape would be used to estimate class, this may cause the model to be over fit and not provide an accurate estimate of unsampled plastral variation. The maximum number of PCs allowed to be used as predictors was \Sexpr{floor(max.ad)} because of both the number of parameters estimated per model and the necessary sample size needed to estimate that many parameters accurately.

@@

\subsection{Machine learning analyses}
\subsubsection{Unsupervised learning}

\subsubsection{Supervised learning}

>>>>

\subsection{Machine learning analyses}
\subsubsection{Unsupervised learning}
% clustering
%   Riemannian shape distance
Because shape space, or configurations after Procrustes superimposition, is a Riemannian manifold \citep{Dryden1998a} the dissimilarity between each landmark configuration was measured as the Riemmanian shape distance \citep{Dryden1998a} AND KENDALL CITATION.

%   PAM
The dissimilarity matrix of shape was divisivly clustering using partioning around mediods (PAM) which is analogous to \textit{k}-means clustering except that instead of minimizing the sum of squared Euclidean distances between observations and centroids, the sum of squared dissimilarities between observations and mediods is minimized CITATION.

%   gap statistic
The optimal number of clusters of shape configurations is unknown being possibly three, four, or some other unknown. Clustering solutions were estimated for between 1 and \Sexpr{nrow(tadult.gap$Tab)} clusters. Clustering solutions were compared using the gap statistic, which is a measure of goodness of clustering CITATION.

%  R implementation
PAM clustering and gap statistic calculation was conducted using the \texttt{cluster} package for R \citep{Maechler2013}.

@@

% model training
%   tuning parameters
%     grid search
%       what are the tuning parameters for each model?
%     best AUC ROC
In order to prevent over fitting each machine learning model, tuning parameters were estimated using 10-fold cross-validation (CV) across a grid search of all tuning parameter combinations. Optimal tuning parameter values were selected based on area under the receiver operating characteristic curve (AUC ROC). Multiclass AUC ROC was estimated using the all-against one strategy derived by \citet{Hand2001} in implemented in PROC PACKGE.

%   model selection
%     AICc
For the multinomial logistic regression models, PCs were added sequentially in order to increase the overall amount of variation in shape included in each model and the final model was that with the lowest AICc \citep{Burnham2002a} AKAIKE AND OTHER CITATION. This procedure was used because the optimal number of PCs to include is unknown, and while including all of the PCs of shape would mean that all of the variability in plastron shape would be used to estimate class, this may cause the model to be over fit and not provide an accurate estimate of unsampled plastral variation. The maximum number of PCs allowed to be used as predictors was \Sexpr{floor(max.ad)} because of both the number of parameters estimated per model and the necessary sample size needed to estimate that many parameters accurately.

@@

\subsection{Machine learning analyses}
\subsubsection{Unsupervised learning}
% figure
%   gap statistic results

\subsubsection{Supervised learning}
% figure
%   ROC model selection (see talk)
%   generalize densities
%     facet: multinomial, rf (see talk)
%   maps
%     facet: training, multinomial, rf
%
% table
%   AICc model selection

### 2013-07-01 documents/turtle_paper.tex
methods. pam.

\subsection{Machine learning analyses}
\subsubsection{Unsupervised learning}
% PAM
%   Riemannian shape distance
%   gap statistic
PAM was conducted using the \texttt{cluster} package for R \citep{Maechler2013}.

@@

% model training
%   tuning parameters
%     grid search
%       what are the tuning parameters for each model?
%     best AUC ROC
In order to prevent over fitting each machine learning model, tuning parameters were estimated using 10-fold cross-validation (CV) across a grid search of all tuning parameter combinations. Optimal tuning parameter values were selected based on area under the receiver operating characteristic curve (AUC ROC). Multiclass AUC ROC was estimated using the all-against one strategy described in \citet{Hand2001}.

%   model selection
%     AICc
For the multinomial logistic regression models, PCs were added sequentially in order to increase the overall amount of variation in shape included in each model and the final model was that with the lowest AICc \cite{Burnham2002a}. This procedure was used because the optimal number of PCs to include is unknown, and while including all of the PCs of shape would mean that all of the variability in plastron shape would be used to estimate class, this may cause the model to be over fit and not provide an accurate estimate of unsampled plastral variation. The maximum number of PCs allowed to be used as predictors was 10 because of both the number of parameters estimated per model and the necessary sample size needed to estimate that many parameters accurately.

>>>>

\subsection{Machine learning analyses}
\subsubsection{Unsupervised learning}
% clustering
%   Riemannian shape distance
Because shape space, or configurations after Procrustes superimposition, is a Riemannian manifold \citep{Dryden1998a} the dissimilarity between each landmark configuration was measured as the Riemmanian shape distance \citep{Dryden1998a} AND KENDALL CITATION.

%   PAM
The dissimilarity matrix of shape was divisivly clustering using partioning around mediods (PAM) which is analogous to \textit{k}-means clustering except that instead of minimizing the sum of squared Euclidean distances between observations and centroids, the sum of squared dissimilarities between observations and mediods is minimized CITATION.

%   gap statistic
The optimal number of clusters of shape configurations is unknown being possibly three, four, or some other unknown. Clustering solutions were estimated for between 1 and 40 clusters. Clustering solutions were compared using the gap statistic, which is a measure of goodness of clustering CITATION.

%  R implementation
PAM clustering and gap statistic calculation was conducted using the \texttt{cluster} package for R \citep{Maechler2013}.

@@

% model training
%   tuning parameters
%     grid search
%       what are the tuning parameters for each model?
%     best AUC ROC
In order to prevent over fitting each machine learning model, tuning parameters were estimated using 10-fold cross-validation (CV) across a grid search of all tuning parameter combinations. Optimal tuning parameter values were selected based on area under the receiver operating characteristic curve (AUC ROC). Multiclass AUC ROC was estimated using the all-against one strategy derived by \citet{Hand2001} in implemented in PROC PACKGE.

%   model selection
%     AICc
For the multinomial logistic regression models, PCs were added sequentially in order to increase the overall amount of variation in shape included in each model and the final model was that with the lowest AICc \citep{Burnham2002a} AKAIKE AND OTHER CITATION. This procedure was used because the optimal number of PCs to include is unknown, and while including all of the PCs of shape would mean that all of the variability in plastron shape would be used to estimate class, this may cause the model to be over fit and not provide an accurate estimate of unsampled plastral variation. The maximum number of PCs allowed to be used as predictors was 10 because of both the number of parameters estimated per model and the necessary sample size needed to estimate that many parameters accurately.
